---
title: "BDA_Assignment_1"
author: "Pratik Bhujade"
date: "01/11/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Question 1
1. Created the dataset, and cleaning it

```{r}
# Importing the tidyR and zoo Library for analysis
library(tidyr)
library(zoo)
# Inserting the dataset
df <- read.csv('DiamondData.csv')
# Removing the spelling mistake in cuts.
df_cleaning <- df
df_cleaning$cut[df_cleaning$cut == "Very Geod"] <-"Very Good"
# Insert means rows in rows with na value
df_clean <- data.frame('cut'= df_cleaning$cut, 'color' = df_cleaning$color, 'clarity' = df_cleaning$clarity)
df_clean$carat <- na.aggregate(df_cleaning$carat)
df_clean$table <- na.aggregate(df_cleaning$table)
df_clean$x <- na.aggregate(df_cleaning$x)
df_clean$y <- na.aggregate(df_cleaning$y)
df_clean$z <- na.aggregate(df_cleaning$z)
df_cleaning$depthh <- (2*df_clean$z/(df_clean$x+df_clean$y))*100
df_clean$depth <- na.aggregate(df_cleaning$depthh)
df_clean$price <- na.aggregate(df_cleaning$price)
df_cleaned <- df_clean
#names(df_cleaned) <- c('cut','color','clarity','carat','table','x','y','z','depth')
summary(df_clean)
```

## Question 2

Summary and Plots

```{r pressure, echo=FALSE}
# Summary Statistics for the whole dataset
print(summary(df_cleaned))
# Pie Chart for different cuts
pie(summary(df_cleaned$cut), main = "Cut - Pie Chart")
# Bar plots for different colors
barplot(summary(df_cleaned$color), xlab = "Colors", ylab = "Frequency", main = "Color - Bar Chart")
# Histogram for different depths
hist(df_cleaned$depth, xlab = "Depths", ylab = "Frequency", main = "Histogram - Depth")
# Scatterplot for carats vs Depth
plot(df_cleaned$carat, df_cleaned$depth, xlab = "Carats", ylab = "Depth", main = "Scatterplot - Carats vs Depth")
```
## Question 3
### Part A
Histogram of price along with the summary Statistics
```{r}
# importing psych library for statistics
library(psych)
# histogram for prices
hist(df_cleaned$price, xlab = "Price", ylab = "Distribution", main = "Histogram of Price")
# Detailed summary statistics from psych library
describe(df_cleaned$price)
```
The Histogram indicates that the distribution is positively skewed, and concentrated to the left. It also indicates that the price is not a normal distribtion, and the higest number of units lie in the distribution that is less than 5000 price.
The summary statistics indicate that mean is 3938.64, and the media is 2401. Moreover the standard deviation is 3994.8.

### Part B
```{r}
# Low range 0-2401, first quartile
price_low <- 2401
# Medium Range from 2402-5345, Third quartile
price_med <- 5345
# High range greater than 5345
low <- subset(df_cleaned, price <= price_low) 
med <- subset(df_cleaned, price <= price_med)
high <- subset(df_cleaned, price > price_med)
# Summary statistics for low, medium and high
summary(low)
summary(med)
summary(high)
```

### Part C
Box plot for different cuts
```{r}
# boxplot for prices, based on different cuts
boxplot(df_cleaned$price~df_cleaned$cut, xlab = "Cuts", ylab = "Prices",main = "Price distribution of cuts")
```

### Part D
```{r}
# creating the dataframe for numerical variables
n_df <- data.frame(df_cleaned$carat,df_cleaned$price,df_cleaned$depth,df_cleaned$table,df_cleaned$x,df_cleaned$y,df_cleaned$z)
# Correlation for numerical variables
cor(n_df)
```

## Question 4
Frequencies and Scatterplots
```{r}
# Importing ggplot for visualization
library(ggplot2)
# SUmmary for different cuts
summary(df_cleaned$cut)
# Summary for different clarities
summary(df_cleaned$clarity)
# Relationship between carat and price on the basis of cuts
ggplot(df_cleaned, aes(price, carat, color = cut)) + geom_point()
# Relationship between carat and price on the basis of clarities
ggplot(df_cleaned, aes(price, carat, color = clarity)) + geom_point()
```
## Question 5
### Part A
```{r}
# creating volume dataframe
volume =data.frame(df_cleaned$x, df_cleaned$y, df_cleaned$z)
# putting names in it
names(volume) <- c('x','y','z')
# calculating volume
volume$volume <- volume$x*volume$y*volume$z
# adding price to the mix
volume$price <- df_cleaned$price
# ploting volume and price
ggplot(volume, aes(volume, price)) + geom_point(color='blue') + geom_smooth(method = 'lm', formula = y ~ poly(x,2))

```
This indicates a polynomial regression, and it indicates that the regression is increasing in magnitude along with being positive.

### Part B
Relationship between carat and volume
```{r}
# adding carats to the mix
volume$carat <- df_cleaned$carat
# plotting volume and carat
ggplot(volume, aes(volume, carat)) + geom_point(color='red') + geom_smooth(method = 'lm', formula = y ~ x)
```
No linear relationship between volum and carats.

### Part C
Relationship between table and depth
```{r}
# Adding table and depth to the mix
volume$table <- df_cleaned$table
volume$depth <- df_cleaned$depth
# ploting table and depth 
ggplot(volume, aes(depth, table)) + geom_point(color='green') + geom_smooth(method = 'lm', formula = y ~ x)
```
Negative relationship between depth and table
### Part D
Relationship of table with all the variables

```{r}
# finding the correlation between all the added variables
cor(volume)
# ploting volume, price and carats with table
ggplot(volume, aes(volume, table)) + geom_point(color='pink') + geom_smooth(method = 'lm', formula = y ~ x)
ggplot(volume, aes(price, table)) + geom_point(color='orange') + geom_smooth(method = 'lm', formula = y ~ x)
ggplot(volume, aes(carat, table)) + geom_point(color='purple') + geom_smooth(method = 'lm', formula = y ~ x)
```


